\vspace{1.0cm} \LARGE \bf MERS-CoV spillover at the camel-human interface

### 2017-11-21 mers-structure.tex
Explain SC logic, per #21

\subsection*{MERS-CoV epidemiology}
%\lmc{Detecting?}
In this study we aimed to understand the drivers of MERS coronavirus transmission in humans and what role the camel reservoir plays in perpetuating the epidemic in the Arabian peninsula by using sequence data collected from both hosts (174 from humans and 100 from camels).
We showed that currently existing models of population structure \citep{vaughan_efficient_2014} can identify distinct demographic modes in MERS-CoV genomic data, where viruses continuously circulating in camels repeatedly jump into humans and cause small outbreaks doomed to extinction (Figures \ref{mcc}, \ref{exploded}).
This inference succeeds under different choices of priors for unknown demographic parameters (Figure \ref{prior}) and in the presence of strong biases in sequence sampling schemes (Figure \ref{mers_epi}).
When rapid coalescence in the human deme is not allowed (Figure \ref{equal_sizes}) structured coalescent inference loses power and ancestral state reconstruction is nearly identical to that of discrete trait analysis (Figure \ref{dta}).

>>>>

\subsection*{MERS-CoV epidemiology}
%\lmc{Detecting?}
In this study we aimed to understand the drivers of MERS coronavirus transmission in humans and what role the camel reservoir plays in perpetuating the epidemic in the Arabian peninsula by using sequence data collected from both hosts (174 from humans and 100 from camels).
We showed that currently existing models of population structure \citep{vaughan_efficient_2014} can identify distinct demographic modes in MERS-CoV genomic data, where viruses continuously circulating in camels repeatedly jump into humans and cause small outbreaks doomed to extinction (Figures \ref{mcc}, \ref{exploded}).
This inference succeeds under different choices of priors for unknown demographic parameters (Figure \ref{prior}) and in the presence of strong biases in sequence sampling schemes (Figure \ref{mers_epi}).
When rapid coalescence in the human deme is not allowed (Figure \ref{equal_sizes}) structured coalescent inference loses power and ancestral state reconstruction is nearly identical to that of discrete trait analysis (Figure \ref{dta}).
When allowed different deme-specific population sizes, the structured coalescent model succeeds because a large proportion of human sequences fall into tightly connected clusters, which informs a low estimate for the population size of the human deme.
This in turn, informs the inferred state of long ancestral branches in the phylogeny, i.e.\ because these long branches are not immediately coalescing, they are most likely in camels.
